The Right Way to Search Evolving Graphs

### 2016-01-29 paper.tex
More comprehensive conclusion

In Section~\ref{sec:breadth-first-search}, we explain how the BFS algorithm can be applied correctly on
evolving graphs. Section~\ref{sec:temporal-paths} provides an example showing that considering
only products of the time-dependent adjacency matrices can lead to incorrect enumeration
of temporal paths. We present and demonstate the BFS algorithm over evolving graphs
in Section~\ref{sec:evolving-graph-bfs}, showing that it is formally equivalent to
BFS over a particular static graph generated by connecting
together active notes of the evolving graph. This static graph generates an algebraic
representation of the BFS as power iteration of its adjacency matrix to a starting
search node, as shown in Section~\ref{sec:bfsla}. The algebraic formulation also
demonstrates interesting connections between properties of the BFS algorithm and
the adjacency matrix. Finally in Section~\ref{sec:applications},
we explain how BFS on evolving graphs may be applied to study dynamical processes
over citation networks.

@@

This paper contemplated one of basic archetypes of evolving graph traversal algorithms. The treatment interweaved graph theory with linear algebra.
Correct tracking of temporal paths on active nodes
provides the right way to search evolving graphs,
which can also be applied to other algorithms for evolving graphs.
% This paper introduced the generalization of BFS to evolving graphs and presented
% an equivalent formulation using linear algebra. The concept of temporal paths
% connecting active nodes turns out to be critical for the correct generalization
% of BFS to evolving graphs. An efficient implementation of BFS on evolving
% graphs is now available as a open source Julia package.
% This paper leaves important extensions open for further research. First, extensions
% of BFS to weighted evolving graphs can provide more valuable information when the degree of connectivity is available in the data.
% Second, efficient algebraic representations and algorithms need to be addressed.
% One major advantage of representing evolving graphs as sparse matrices
% is the matrix representation is highly structured so it is easy to design
% high performance parallel implementation.
% Third, depth-first search and other graph traversal algorithms can be
% extended using similar ideas.

>>>>

In Section~\ref{sec:breadth-first-search}, we explain how the BFS algorithm can be applied correctly on
evolving graphs. Section~\ref{sec:temporal-paths} provides an example showing that considering
only products of the time-dependent adjacency matrices can lead to incorrect enumeration
of temporal paths. We present and demonstrate the BFS algorithm over evolving graphs
in Section~\ref{sec:evolving-graph-bfs}, showing that it is formally equivalent to
BFS over a particular static graph generated by connecting
together active notes of the evolving graph. This static graph generates an algebraic
representation of the BFS as power iteration of its adjacency matrix to a starting
search node, as shown in Section~\ref{sec:bfsla}. The algebraic formulation also
demonstrates interesting connections between properties of the BFS algorithm and
the adjacency matrix. Finally in Section~\ref{sec:applications},
we explain how BFS on evolving graphs may be applied to study dynamical processes
over citation networks.

@@

The correct generalization of BFS to evolving graphs necessitates a careful
enumeration of temporal paths which is not always captured by products of
successive adjacency matrices from the snapshots. BFS over any evolving graph $G_n$
is formally equivalent to BFS on a related static graph $G$ formed by connecting only
active nodes across different times. The adjacency matrix $\bm A_n$ of $G$ correctly
encodes causality of the temporal paths. When specialized to acyclic $G_n$, the matrix
$\bm A_n$ is also nilpotent, thus providing a simple proof that BFS terminates.
Nevertheless, the BFS can be computed efficiently without ever explicitly constructing
$\bm A_n$, as only its action on a vector is required for the power iterations.
The new concepts of activeness, temporal paths, and the explicit construction of
an equivalent static graph $G$ that make possible a correct implementation of BFS
to evolving graphs provide powerful new insignts into how similar graphical
algorithms may be generalized correctly.

% This paper introduced the generalization of BFS to evolving graphs and presented
% an equivalent formulation using linear algebra. The concept of temporal paths
% connecting active nodes turns out to be critical for the correct generalization
% of BFS to evolving graphs. An efficient implementation of BFS on evolving
% graphs is now available as a open source Julia package.
% This paper leaves important extensions open for further research. First, extensions
% of BFS to weighted evolving graphs can provide more valuable information when the degree of connectivity is available in the data.
% Second, efficient algebraic representations and algorithms need to be addressed.
% One major advantage of representing evolving graphs as sparse matrices
% is the matrix representation is highly structured so it is easy to design
% high performance parallel implementation.
% Third, depth-first search and other graph traversal algorithms can be
% extended using similar ideas.
